RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-07-04 background.tex
Section 1, 2 and the example

@@

Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}). An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Tycoon}, Figure~\ref{fig:example-small}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \texttt{Show Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}).

Different links connect the elements in a GRL model. AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. The goal \texttt{Generate Cars} depends on the goal \texttt{Value Has Changed}. %SG: I added it to the model but we can also remove it altogether. 
%\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}

@@

\begin{itemize}
	\item Why does actor \texttt{Traffic Tycoon} have only a single softgoal \texttt{Show Simulation}? Why is this, for instance, not connected to goal \texttt{Simulate}? %SG: I changed this question but let me know if you want to change them back and then change the graph.
	\item What does \texttt{Keep same cars} mean?
	\item Why does task \texttt{Keep same cars} contribute positively to \texttt{Static simulation} and negatively to \texttt{Dynamic simulation}?
	\item Why does \texttt{Generate Cars} XOR-decompose into two tasks?
\end{itemize}

These are the type of the questions that we cannot answer by just looking at GRL models. The model in Figure~\ref{fig:Example1} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

\begin{itemize}
\item[] We have goal $G$,
\item[] Doing action $A$ realizes to goal $G$,
\item[] Which will contribute positively to the softgoal $S$
\item[] \textit{Therefore} 
\item[] We should perform action $A$
\end{itemize}

The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals. Once instantiated, we obtain an argument. We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows: %SG: Need to update the example.
\begin{itemize}
\item[] We have goal \texttt{Static simulation},
\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
\item[] Which contribute positively to the softgoal \texttt{Simulate} 
\item[] \textit{Therefore} 
\item[] We should perform action \texttt{Keep same cars}.
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument. %SG:placeholder for the examples.

Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3). For example,  can argue that another goal \texttt{Semi-dynamic Simulation} also realizes the softgoal \texttt{Simulate}.
\todo{Marc}{Sepideh}{Do you think our example is good? I think it is confusing. Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader. Now the tasks are all a bit vague and I think nobody knows what they actually mean.}

>>>>

In this Section and Section~\ref{}, we focus on the smaller example of traffic simulator case. In Section~\ref{} we provide the detailed example of the traffic simulator and its model.

@@

Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}). An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \texttt{Improve Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Simulate Traffic}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars} which are alternative way of achieving the goal, \texttt{Generate Cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).

Different links connect the elements in a GRL model. AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. Here, the goal \texttt{Generate Cars} depends on the resource \texttt{Car Objects}. %SG: I added it to the model but we can also remove it altogether. 
%\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}

@@

\begin{itemize}
	\item Why does actor \texttt{Traffic Simulator} have softgoal \texttt{High Performance}which is not linked to any of the goals \texttt{Generate Cars} and \texttt{Simulate Traffic}? %SG: I changed this question but let me know if you want to change them back and then change the graph.
	\item What does \texttt{Keep same cars} mean?
	\item Why does task \texttt{Create New cars} contribute negatively to \texttt{Static simulation} and positively to \texttt{Dynamic simulation}?
	\item Why does \texttt{Simulate Traffic} AND-decompose into two tasks?
\end{itemize}

These are the type of the questions that we cannot answer by just looking at GRL models. The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

\todo{Sepideh}{Marc}{Read this part and see if it makes sense.}
\begin{itemize}
\item[] We have goal $G$,
\item[] Doing action $A$ realizes to goal $G$,
\item[] Which will contribute positively to the softgoal $S$
\item[] \textit{Therefore} 
\item[] We should perform action $A$
\end{itemize}

The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals. Once instantiated, we obtain an argument. We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows: 
\begin{itemize}
\item[] We have goal \texttt{Simulate Traffic},
\item[] Doing action \texttt{Dynamic Simulation} realizes goal \texttt{Simulate Traffic}, 
\item[] Which contribute positively to the softgoal \texttt{Improve Simulation} 
\item[] \textit{Therefore} 
\item[] We should perform action \texttt{Dynamic Simulation}.
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 1: if we find that  \texttt{Dynamic Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.

Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3). For example,  can argue that another task \texttt{Semi-dynamic Simulation} also realizes the goal \texttt{Simulate Traffic}.
%\todo{Marc}{Sepideh}{Do you think our example is good? I think it is confusing. Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader. Now the tasks are all a bit vague and I think nobody knows what they actually mean.}
